The Neuro Bureau ADHD-200 Preprocessed Repository

### 2015-12-19 Bellec_ADHD200_2016.tex
Fixed methods and citations.

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%%
%%
%% $Id: elsarticle-template-num.tex 4 2009-10-24 08:22:58Z rishi $
%%
%%
\documentclass[preprint,12pt,1p]{elsarticle}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}

@@

\title{The ADHD200 preprocessed experiment:\\lowering the barriers to a neuroimaging competition}

%\tnotetext[label0]{This is only an example}

\author[label0,label1,label2]{Pierre Bellec\corref{cor1}}
\address[label0]{The Neuro Bureau Research Institute, Leipzig, Germany}
\address[label1]{Centre de Recherche de l'Institut Universitaire de G\'eriatrie de Montr\'eal, Montr\'eal, CA}
\address[label2]{D\'epartement d'Informatique et de Recherche Op\'erationnelle, Universit\'e de Montr\'eal, Montr\'eal, CA}

@@

\author[label0,label5]{Daniel S. Margulies}
\address[label5]{Max Planck Insitute, Leipzig, GER}
5\ead{daniel.margulies@gmail.com}

@@

\begin{abstract}
Text of abstract. Text of abstract. Text of abstract. Text of abstract. Text of abstract. 
\end{abstract}

@@

%%
%% Start line numbering here if you want
%%
% \linenumbers

\section{Introduction}

In 2011, the ``ADHD-200 Global Competition'' was held with the aim of engaging researchers from a variety of analytical backgrounds to identify biomarkers of attention deficit hyperactivity disorder (ADHD) from resting-state functional magnetic resonance imaging (R-fMRI) data \cite{Milham2012}. The competition made use of the ``ADHD-200 Sample'' data collection that was aggregated from eight independent studies and shared through the Intenational Neuroimaging Datasharing Initiative (INDI) \cite{Mennes2013}. The data includes R-fMRI, structural MRI (sMRI), and basic phenotypic information for 973 individuals (554 TDC, 419 ADHD) \cite{Milham2012}. Competitors were given five and a half months to optimize a classification algorithm on training data (776 individuals) and submit their predicted clinical labels for test data for which diagnostic information was withheld. The competition data was distributed in a raw form and before any analysis could begin, had to be preprocessed to make it comparable across individuals and to remove nuisance variation that would obscure the underlying biological signal. These preprocessing steps present a significant hurdle for would-be competitors who do not have the specialist knowledge of neuroimaging methods or access to the high performance computing resources required to process such a large data collection in a short amount of time. Realizing this barrier to entry for  statisticians and computer scientists, who were potentially the most qualified for the machine learning aspect of the competition, the Neuro Bureau prospectively collaborated with all competitors by preprocessing the data and sharing these results.

The ``ADHD-200 Preprocessed'' is a repository of preprocessed R-fMRI and sMRI data along with statistical derivatives from the ADHD-200 Sample. Rather than favoring a specific processing strategy, we followed a pluralistic approach by preprocessing the data using multiple pipelines (called ``Athena'', ``Burner'' and ``NIAK'') that differed in toolsets used, the philosophy motivating choices of algorithms and parameters, and the statistical derivatives calculated. The Athena pipeline processing R-fMRI and sMRI images using a custom BASH script that combines functionality from AFNI \cite{Cox1996} and FSL \cite{Smith2004} neuroimaging toolkits. The Burner pipeline used SPM8 \cite{ashburner} to process sMRI data for voxel-based morphometry style analyses. The NIAK pipeline was implemented using tools from the neuroimaging analysis kit (NIAK) \cite{bellecNIAK} pipelined together using the Pipeline System for Octave and Matlab (PSOM) \cite{slcPSOM} to process both sMRI and R-fMRI datasets. The results from the pipelines are shared on the Neuro Bureau project page at the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC)\footnote{\url{http://www.nitrc.org/frs/?group_id=383}}. The data was used by several teams who competed in the ADHD-200 Global Competition, including the winning entry by a group of biostaticians with limited prior neuroimaging experience \cite{Eloyan2012}. The initiative has had an significant impact beyond the competition, having been used in 49 publications \cite{Rangarajan2014, Liang2012, Tabas2014, Rangarajan2015, Mahanand2013, Lifshitz2012, Fujita2013, Ji2011, Li2015, Li2013, Liu2012, DosSantosSiqueira2014, Olivetti2014, Han2015, Wang2013a, Subramanian2013, Dey2014, Bellec2012, Bohland2012, Chang2012, Cheng2012, Colby2012, Dai2012, Dey2012, Eloyan2012, Olivetti2012, Sato2012a, Carmona2015, Carmona2015a, Hou2015, Deshpande2015, She2014, Lavoie-Courchesne2012b, Chen2015, Nunez-Garcia2015, Solmaz2012, Anderson2014, KadkhodaeianBakhtiari2012, Sato2013, Kyeong2015, Sato2012, Takahashi2012, He2013, Kong2013, Yao2013, Yang2015, Ahn2015, Fujita2014, Reiss2014}, three PhD theses \cite{Colby2012a, Dey2013, Zhang2012}, three master's dissertations \cite{VanGalenLast2011, Vidal2014, Wang2013} and one patent \cite{Dey2013} in just over 3 years, with more in press or under review.

@@

The ADHD-200 preprocessed data was released in the summer of 2012 and can be downloaded from NITRC\footnote{\url{http://www.nitrc.org/frs/?group_id=383}}. No data usage agreement is required to access or download the data, the only requirement is registering for a free NITRC account. This registration enables downloads to be tracked for usage statistics and so that users can be contacted in the event that errors are found. The ADHD-200 Sample allows unrestricted data usage for non-commercial research purposes provided that the specific datasets included in an analysis be cited appropriately and that their funding sources be acknowledged\footnote{\url{http://fcon_1000.projects.nitrc.org/indi/adhd200/}}. There are no more restrictions placed on the preprocessed data or derivatives other than the request that the ADHD-200 Preprocessed Initiative is cited appropriately and that the specific pipeline is acknowledged in publications using the data. A forum is available on the Neuro Bureau's NITRC project page for users to ask questions or report
problems\footnote{\url{http://www.nitrc.org/forum/forum.php?forum_id=2046}}. Questions regarding data acquisition or phenotypic variables should be directed to INDI's support forum\footnote{\url{http://www.nitrc.org/forum/forum.php?forum_id=1735}}. 
\par
ADHD-200 Preprocessed data is available as a collection of compressed tar files that are organized by pipeline, sites of data collection, training and test samples, as well as by derivative. The following preprocessed derivatives are available.

\begin{table}[!ht]
\caption{{\bf Summary of ADHD-200 data by site.}}\label{part_tab}
  \begin{tabular}{L{.1in}L{.1in}L{1.1in}L{1.1in}L{1.1in}L{.7in}}
    \hline
	\\
    \multicolumn{3}{l}{\textbf{Bradley Hostpital/Brown University}} & & \\
    & \multicolumn{4}{l}{\emph{Scanner:} Siemens TIM Trio 3T}\\
    & \multicolumn{5}{l}{\emph{sMRI:} 3D MPRAGE, FA = 9$^{\circ}$, TR = 2250 ms, TE = 2.98 ms,}\\
    & &  \multicolumn{3}{l}{TI = 900 ms, $1\times1\times1$ mm$^3$} \\
    & \multicolumn{4}{l}{\emph{R-fMRI:} EPI w/ prospective motion correction (PACE), FA = 9$^{\circ}$,}\\
    & & \multicolumn{4}{l}{TR = 2000 ms, TE = 25 ms, 35 3-mm thick slices, $3.0 \times 3.0$ mm$^2$}\\
    & & \multicolumn{4}{l}{in-plan resolution. Participants were asked to relax with their eyes}\\
    & & \multicolumn{4}{l}{open and fixate on the text ``relax'' that was projected on a screen.}\\
    & \multicolumn{2}{l}{\emph{TDC}:} 27, 
    %& \multicolumn{4}{l} \emph{T1 weighted anatomical scans:}  
	\hline
	\end{tabular}
\end{table}

\subsection{The Athena and NIAK pipelines}

\paragraph{Preprocessed sMRI and R-fMRI} Both the Athena and NIAK pipelines released denoised R-fMRI volumes resampled into a reference, stereotaxic space (variants of the MNI template) as compressed 4D nifti files (.nii.gz). Transformations from native to stereotaxic space for each dataset were released in separate files, in FSL .xfm transform files for Athena and in .xfm MINC files for NIAK. In addition, the estimated motion parameters (bundled with additional intermediate outputs) were released in separate archives, inside text files (AFNI .1D tab-separated-value files) for Athena, and inside Matlab HDF5 files for NIAK.

\paragraph{Regional time series} Regional time series were extracted based on a series of different brain parcellations. The Athena ppipeline used functional parcels generated by a spectral clustering method (Craddock et al. 2012) as well as a number of parcellations established from anatomical landmarks, with a resolution ranging from 93 to 351 parcels. The NIAK pipeline only used functional parcels generated by a region growing method (Bellec et al. 2006) with two resolutions: 954 and 2843 parcels. The Athena pipeline distributed regional time series as text files (.1D AFNI) and the NIAK distributed them as HDF5 files (.mat Matlab). All brain parcellations were also released as compressed nifti files (.nii.gz). Note that the phenotypic information of the ADHD200 sample was also released as comma-separated-values files (.csv).

\paragraph{Other derivatives}
The Athena pipeline derived a number of additional metrics from the preprocessed fMRI datasets: fractional amplitude of low frequency fluctuations (fALFF) maps, regional homogeneity (ReHO) maps and dual-regression based on the (Smith et al. 2009) group independent component analysis (ICA) with 10 resting-state networks. One compressed nifti file (.nii.gz) was released per fMRI dataset and type of derivative, with in addition the time courses of the dual regression available in text files (.1D AFNI).

\subsection{The Burner pipeline}

@@

\section{Processing strategies}

\subsection{Preprocessed sMRI and R-fMRI}
% http://onlinelibrary.wiley.com/doi/10.1002/hbm.22790/epdf
\paragraph{The Athena pipeline} The preprocessing steps of functional MRI datasets included removal of the first four volumes to allow for magnetization to reach equilibrium, slice timing correction, motion correction, normalization to MNI152 standard space at 4 mm isotropic voxel resolution, regression of nuisance covariates including head motion parameters, global mean, white matter, and cerebrospinal ﬂuid signals, band-pass filtering of fMRI ($0.009 Hz < f < 0.08 Hz$) to reduce low-frequency drift and high-frequency physiological noise, and finally spatial smoothing with a 6 mm full-width-at-half-maximum isotropic Gaussian kernel.

\paragraph{The NIAK pipeline} The preprocessing steps followed those of the Athena pipeline with the following differences: (1) only the first three volumes were suppressed, instead of four; (2) only a high-pass filter was implemented ($0.009 Hz < f$), instead of a band pass, and no regression of confounds such as motion parameters, global average, etc. Instead, a physiological noise correction based on automatic labeling of ICA components was implemented \citep{Perlbarg2007}. The fMRI volumes were resampled in stereotaxic space at 3 mm isotropic (instead of 4 mm), and this was done only right before spatial smoothing. Finally note that although both Athena and NIAK used the MNI non-linear stereotaxic template space described in \citep{Fonov2011}, the NIAK used the young adult (ICBM152) symmetric variant, while Athena used the NIHPD asymmetric pediatric template (4.5 y.o.-18.5 y.o.).

\subsection{Regional time series}

For both Athena and NIAK, regional time series were extracted by averaging preprocessed resting-state volumes within different brain regions. The Athena pipeline first used a series of anatomical templates:
\begin{itemize}
\item The automated anatomical labeling (AAL) template \citep{Tzourio-Mazoyer} 
\item The Eickhoff-Zilles (EZ) [8]
\end{itemize}
, , Talairach and Tournoux (TT) [9], and Harvard-Oxford (HO) anatomical atlases. In addition,  as well as, 200 (CC200) and 400 (CC400) ROI atlases derived from functionally parcellating the resting state data [10]. The AAL atlas is distributed with the AAL Toolbox. The atlas was fractionated into functional space using nearest-neighbor interpolation.
The EZ atlas was derived from the max-propagation atlas distributed with the SPM Anatomy Toolbox. The atlas was transformed into template space using the Colin 27 template (also distributed with the toolbox) as an intermediary and fractionated into functional space using nearest-neighbor interpolation.
The HO atlas distributed with FSL is split into cortical and subcortical probabilistic atlases. A 25\% threshold was applied to each of these atlases and they were subsequently bisected into left and right hemispheres at the midline (x=0). ROIs representing left/right WM, left/right GM, left/right CSF and brainstem were removed from the subcortical atlas. The subcortical and cortical ROIs were combined and then fractionated into functional space using nearest-neighbor interpolation.
The TT atlas distributed with AFNI was coregistered and warped into template space and subsequently fractionated into functional space using nearest neighbor interpolation.
\par
Functional parcellation was accomplished using the two-stage spatially-constrained functional procedure described in [10]. Preprocessed and unfiltered resting state data corresponding to 650 subjects with the cleanest data (as determined by registration quality and applying a max translation < 3mm and max rotation < 3 degrees). A grey matter mask was constructed for the subjects by averaging grey matter masks derived from freesurfer automated segmentation. Subject specific connectivity graphs were constructed by treating each withen-gm-mask voxel as a node and edges corresponding to super-threshold temporal correlations to the voxels' 3D (27 voxel) neighborhood. Each subjects' graph was partitioned into 200 or 400 regions using normalized cut spectral clustering. Association matrices were constructed from the clustering results by setting the connectivity between voxels to 1 if they are in the same ROI and 0 otherwise. A group-level correspondance matrix was constructed by averaging the individual level association matrices and subsequently partitioned into 200 or 400 regions using normalized cut clustering.
\par 
Regional homogeneity was calculated for the preprocessed and filtered whole brain resting state data using the procedure described in [11]. The scripts used to calculate ReHo were generously provided by Dr. Xinian Zuo. 
\par 
The FC maps for the Smith PNAS 2009 RSNs were constructed using a modified dual regression approach. I used a spatial multiple regression to extract time courses corresponding to spatial templates of the 10 Smith PNAS 2009 RSNs, and then independently correlated each of the 10 time courses with whole brain data to generate subject-specific FC maps for each RSN. Dr. Steve Smith suggested that instead of using univariate correlation to generate the FC maps, that I use multiple regression instead. These files are the result of that analysis.

\subsection{The Burner Pipeline}
Details on the Burner pipeline can be found on the NITRC website\footnote{\url{http://www.nitrc.org/plugins/mwiki/index.php/neurobureau:NiakPipeline}}.
T1 images were segmented into grey matter and white matter probability maps using "New Segmentation" in SPM8. The tissue maps were rigidly aligned (translation+rotation). Inter subject normalization was performed using DARTEL toolbox in SPM8. Images were iteratively registered to the group average (population template), and template was updated iteratively. This resulted finer and finer registration. The registration parameters were applied to each image to transform each image into the space of population average. Modulation was applied to conserve the global tissue volumes after normalization.

\subsection{The NIAK pipeline}
Details on the NIAK pipeline can be found on the NITRC website.

>>>>

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%%
%%
%% $Id: elsarticle-template-num.tex 4 2009-10-24 08:22:58Z rishi $
%%
%%
\documentclass[preprint,12pt,3p]{elsarticle}
\usepackage{rotating}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\arraybackslash\hspace{0pt}}p{#1}}

@@

\title{The Neuro Bureau ADHD-200 Preprocessed Repository}

% - CC - What about just sticking to the name of the repository? The experiment
% part might make people think that is not an ongoing data sharing initiative.
% We are sharing more that just the statistical derivatives and
% we are also sharing processed structural MRI. I think that capturing all of that
% in a subtitle would be too long.
% - PB - Agreed, it's a cool title.

% The ADHD-200 Preprocessed experiment: lowering the barriers to a neuroimaging competition
% The ADHD-200 Preprocessed Repository:\Openly sharing preprocessed data and derivatives

\author[label0,label1,label2]{Pierre Bellec\corref{cor1}}
\address[label0]{The Neuro Bureau}
\address[label1]{Centre de Recherche de l'Institut Universitaire de G\'eriatrie de Montr\'eal, Montr\'eal, CA}
\address[label2]{D\'epartement d'Informatique et de Recherche Op\'erationnelle, Universit\'e de Montr\'eal, Montr\'eal, CA}

@@

\author[label0,label5]{Daniel S. Margulies}
\address[label5]{Max Planck Research Group for Neuroanatomy \& Connectivity, Max Planck Institute for Human Cognitive and Brain Sciences, Leipzig, Germany}
\ead{daniel.margulies@gmail.com}

@@

\begin{abstract}
In 2011, the ``ADHD-200 Global Competition'' was held with the aim of identifying biomarkers of attention deficit hyperactivity disorder from resting-state functional magnetic resonance imaging (rs-fMRI) and structural MRI (s-MRI) data collected on 973 individuals. Statisticians and computer scientists were potentially the most qualified for the machine learning aspect of the competition, but generally lacked the specialized skills to implement the necessary steps of data preparation for rs-fMRI. Realizing this barrier to entry, the Neuro Bureau prospectively collaborated with all competitors by preprocessing the data and sharing these results at the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC) (\url{http://www.nitrc.org/frs/?group_id=383}). This ``ADHD200 preprocessed'' release included multiple analytical pipelines to cater to different philosophies of data analysis. The processed derivatives included denoised and registered 4D fMRI volumes, regional time series extracted from brain parcellations, maps of 10 intrinsic connectivity networks, fractional amplitude of low frequency fluctuation, and regional homogeneity, along with grey matter density maps. The data was used by several teams who competed in the ADHD-200 Global Competition, including the winning entry by a group of biostaticians. To the best of our knowledge, the ADHD200 preprocessed release was the first large public resource of preprocessed resting-state fMRI and structural MRI data, and remains to this day the only resource featuring a battery of alternative processing paths. 
\end{abstract}

@@

%%
%% Start line numbering here if you want
%%
% \linenumbers
\section{Highlights}
\begin{itemize}
\item Processed derivatives generated from structural and resting-state fMRI in 973 individuals. 
\item The sample includes 585 typically developing children, 362 children with attention deficit hyperactivity disorder, and 26 with unknown diagnostic status.
\item Functional MRI volumes were denoised and aligned using two alternative processing pipelines. 
\item Derivatives include regional time series, grey matter segmentation as well as two variants of connectivity maps and a measure of intrinsic brain activity. 
\item These resources are freely available, with minimal access requirements. 
\end{itemize}
\section{Introduction}

In 2011, the ``ADHD-200 Global Competition'' was held with the aim of engaging researchers from a variety of analytical backgrounds to identify biomarkers of attention deficit hyperactivity disorder (ADHD) from resting-state functional magnetic resonance imaging (rs-fMRI) and structural MRI (s-MRI) data \cite{Milham2012}. The competition made use of the ``ADHD-200 Sample'' data collection that was aggregated from eight independent sites and shared through the Intenational Neuroimaging Datasharing Initiative (INDI) \cite{Mennes2013}. The data includes rs-fMRI, structural MRI (sMRI), and basic phenotypic information for 973 individuals (585 TDC, 362 ADHD, 26 unknown) \cite{Milham2012}. Competitors were given five and a half months to optimize a classification algorithm on training data (776 individuals) and submit their predicted clinical labels for test data for which diagnostic information was withheld. The competition data was distributed in a raw form and before any analysis could begin, had to be preprocessed to make it comparable across individuals and to remove nuisance variation that would obscure the underlying biological signal. These preprocessing steps present a significant hurdle for would-be competitors who do not have the specialist knowledge of neuroimaging methods or access to the high performance computing resources required to process such a large data collection in a short amount of time. This category includes statisticians and computer scientists, who were potentially the most qualified for the machine learning aspect of the competition. Realizing this barrier to entry, the Neuro Bureau prospectively collaborated with all competitors by preprocessing the data and sharing these results.

The ``ADHD-200 Preprocessed'' is a repository of preprocessed rs-fMRI and s-MRI data along with statistical derivatives from the ADHD-200 Sample. Rather than favoring a specific processing strategy, we followed a pluralistic approach by preprocessing the data using multiple pipelines (called ``Athena'', ``Burner'' and ``NIAK'') that differed in toolsets used, the philosophy motivating choices of algorithms and parameters, and the statistical derivatives calculated. The Athena pipeline processed rs-fMRI and s-MRI images using a custom Bash script that combined AFNI \cite{cox1996afni} and FSL \cite{smith2004advances} neuroimaging toolkits. The Burner pipeline used SPM8 \cite{ashburner2012spm8} to process s-MRI data for voxel-based morphometry style tools. The NIAK pipeline processed rs-fMRI and s-MRI using tools from the neuroimaging analysis kit (NIAK) \cite{Bellec2011} assembled together into a workflow using the Pipeline System for Octave and Matlab (PSOM) \cite{Bellec2012}. The results from the pipelines are shared on the Neuro Bureau project page at the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC)\footnote{\url{http://www.nitrc.org/frs/?group_id=383}}. The data was used by several teams who competed in the ADHD-200 Global Competition, including the winning entry by a group of biostaticians with limited prior neuroimaging experience \cite{Eloyan2012}. The initiative has had an significant impact beyond the competition, having been used in 49 publications \cite{Rangarajan2014, Liang2012, Tabas2014, Rangarajan2015, Mahanand2013, Lifshitz2012, Fujita2013, Ji2011, Li2015, Li2013, Liu2012, DosSantosSiqueira2014, Olivetti2014, Han2015, Wang2013a, Subramanian2013, Dey2014, Bellec2012, Bohland2012, Chang2012, Cheng2012, Colby2012, Dai2012, Dey2012, Eloyan2012, Olivetti2012, Sato2012a, Carmona2015, Carmona2015a, Hou2015, Deshpande2015, She2014, Lavoie-Courchesne2012b, Chen2015, Nunez-Garcia2015, Solmaz2012, Anderson2014, KadkhodaeianBakhtiari2012, Sato2013, Kyeong2015, Sato2012, Takahashi2012, He2013, Kong2013, Yao2013, Yang2015, Ahn2015, Fujita2014, Reiss2014}, three PhD theses \cite{Colby2012a, Dey2013, Zhang2012}, three master's dissertations \cite{VanGalenLast2011, Vidal2014, Wang2013} and one patent \cite{Dey2013} in just over 3 years, with more in press or under review.

\section{Organization and access to the repository}

The ADHD-200 preprocessed data was released in the summer of 2012 and can be downloaded from NITRC\footnote{\url{http://www.nitrc.org/frs/?group_id=383}}. No data usage agreement is required to access or download the data, the only requirement is registering for a free NITRC account. This registration enables downloads to be tracked for usage statistics and so that users can be contacted in the event that errors are found. The ADHD-200 Sample allows unrestricted data usage for non-commercial research purposes provided that the specific datasets included in an analysis be cited appropriately and that their funding sources be acknowledged\footnote{\url{http://fcon_1000.projects.nitrc.org/indi/adhd200/}}. There are no more restrictions placed on the preprocessed data or derivatives other than the request that the ADHD-200 Preprocessed Initiative is cited appropriately and that the specific pipeline is acknowledged in publications using the data. A forum is available on the Neuro Bureau's NITRC project page for users to ask questions or report
problems\footnote{\url{http://www.nitrc.org/forum/forum.php?forum_id=2046}}. Questions regarding data acquisition or phenotypic variables should be directed to INDI's support forum\footnote{\url{http://www.nitrc.org/forum/forum.php?forum_id=1735}}.

@@

ADHD-200 Preprocessed data is available as a collection of compressed tar files that are organized by pipeline, sites of data collection, training and test samples, as well as by derivative. The following preprocessed derivatives are available.

\subsection{Preprocessed s-MRI and rs-fMRI} Both the Athena and NIAK pipelines released denoised rs-fMRI volumes resampled into a reference, stereotaxic space (variants of the MNI template) as compressed 4D nifti files (.nii.gz). Transformations from native to stereotaxic space for each dataset were released in separate files, in FSL .xfm transform files for Athena and in .xfm MINC files for NIAK. In addition, the estimated motion parameters (bundled with additional intermediate outputs) were released in separate archives, inside text files (AFNI .1D tab-separated-value files) for Athena, and inside Matlab HDF5 files for NIAK.

\subsection{Regional time series} Regional time series were extracted based on a series of different brain parcellations. The Athena ppipeline used functional parcels generated by a spectral clustering method (Craddock et al. 2012) as well as a number of parcellations established from anatomical landmarks, with a resolution ranging from 93 to 351 parcels. The NIAK pipeline only used functional parcels generated by a region growing method (Bellec et al. 2006) with two resolutions: 954 and 2843 parcels. The Athena pipeline distributed regional time series as text files (.1D AFNI) and the NIAK distributed them as HDF5 files (.mat Matlab). All brain parcellations were also released as compressed nifti files (.nii.gz). Note that the phenotypic information of the ADHD200 sample was also released as comma-separated-values files (.csv).

\subsection{Connectivity maps}
The Athena pipeline derived a number of brain maps capturing some aspects of RfMRI connectivity, derived from the preprocessed fMRI datasets: fractional amplitude of low frequency fluctuations (fALFF) maps, regional homogeneity (ReHO) maps and dual-regression based on the (Smith et al. 2009) group independent component analysis (ICA) with 10 resting-state networks. One compressed nifti file (.nii.gz) was released per fMRI dataset and type of derivative, with in addition the time courses of the dual regression available in text files (.1D AFNI).

\subsection{Voxel-based morphometry}

@@

\begin{table}[!ht]
\caption{{\bf ADHD-200 participants by site.} BHBU: Bradley Hospital/ Brown University, KKI: Kennedy Krieger Institute, NI: NeuroIMAGE sample, NYU: New York University Child Study Center, OHSU: Oregon Health Sciences University, PKU: Peking University, Pitt: University of Pittsburgh, WUSTL: Washington University at Saint Louis, avg.: average. $^*$Diagnostic labels are current not available for BHBU, they have been listed as TDC in the table, but not included in the totals.}\label{part_table}
  \begin{tabular}{llrrrr}
      \hline
       & & \multicolumn{2}{c}{TDC} & \multicolumn{2}{c}{ADHD}\\
      Site & Sex & N & Age Range (avg.) & N & Age Range (avg.) \\
        \hline
    \noalign{\vskip 1ex}  
    BHBU & F & 17$^*$ & 8.50 - 17.87 (13.79) 
             & 0 & - \\
         & M &  9$^*$ & 12.00 - 17.83 (15.96) 
             & 0 & - \\
    KKI  & F & 28 & 8.12 - 12.36 (10.19) 
             & 10 & 8.10 - 12.99 (9.97) \\
         & M & 41 & 8.02 - 12.87 (10.35) 
             & 15 & 8.19 - 12.98 (10.08) \\
    NI  & F & 25 & 11.66 - 26.31 (18.69) 
             & 5 & 11.85 - 19.66 (15.07) \\
         & M & 12 & 12.84 - 25.04 (18.02) 
             & 31 & 11.05 - 20.89 (17.05) \\
    NYU  & F & 55 & 7.17 - 17.96 (12.20) 
             & 34 & 7.35 - 17.15 (10.27) \\
         & M & 56 & 7.19 - 17.96 (12.20) 
             & 117 & 7.24 - 17.61 (11.19) \\
    OHSU  & F & 40 & 7.33 - 12.50 (8.98) 
             & 13 & 7.42 - 11.33 (9.01) \\
         & M & 30 & 7.17 - 11.92 (9.48) 
             & 30 & 7.42 - 11.83 (8.94) \\
    PKU  & F & 59 & 8.33 - 15.17 (10.90) 
             & 10 & 8.83 - 15.92 (10.88) \\
         & M & 84 & 8.08 - 14.92 (11.80)  
             & 92 & 8.33 - 17.33 (12.22) \\ 
    Pitt  & F & 44 & 10.16 - 20.45 (15.65) 
             & 1 & 15.04 \\
         & M & 50 & 10.11 - 18.96 (14.55)  
             & 3 & 13.72 - 17.03 (15.55) \\ 
    WUSTL  & F & 28 & 7.09 - 21.81 (11.33) 
             & 0 & - \\
         & M & 33 & 7.25 - 21.83 (11.58)  
             & 0 & - \\ 
    \bf{Totals} & F & 279$^*$ & 7.09 - 26.31 (11.9)
               & 73 & 7.35 - 19.66 (10.48) \\
           & M & 306$^*$ & 7.17 - 25.04 (12.20)
               & 288 & 7.24 - 20.89 (11.90) \\
    \noalign{\vskip 1ex}  
    \hline
	\end{tabular}
\end{table}

\begin{sidewaystable}[!ht]
\caption{{\bf Structural MRI acquisition parameters by site.} Seq: imaging sequence, FA: flip angle, TE: echo time, TR: repetition time, TI: inversion recovery delay, PA: parallel acquisition, Res: voxel resolution, BHBU: Bradley Hospital/ Brown University, KKI: Kennedy Krieger Institute, NI: NeuroIMAGE sample, NYU: New York University Child Study Center, OHSU: Oregon Health Sciences University, PKU: Peking University, Pitt: University of Pittsburgh, WUSTL: Washington University at Saint Louis, Trio: Siemens TIM Trio 3T, Allegra: Siemens Allegra, Avanto: Siemens Avanto, MPRAGE: magnetization prepared rapid gradient echo, S: sensitivity encoding (SENSE), G: generalized auto-calibrating partially parallel acquisition (GRAPPA)}\label{smri_params_table}
  \begin{tabular}{lllllllll}
      \hline
    Site & Scanner & Seq & FA & TE & TR & TI & PA & Res. \\
    \hline
    \noalign{\vskip 1ex}  
    BHBU & Trio 3T & 3D MPRAGE & 9$^{\circ}$ & 2.98 ms & 2250 ms & 900 ms & None & $1.00\times1.00\times1.00$ mm$^3$ \\
    KKI & Phillips 3T & 3D MPRAGE & 8$^{\circ}$ & 3.7 ms & 3500 ms & 1000 ms & S $\times$2 & $1.00\times1.00\times1.00$ mm$^3$ \\
    NI & Avanto 1.5T & 3D MPRAGE & 7$^{\circ}$ & 2.95 ms & 2730 ms & 1000 ms & G $\times$2 & $1.00\times1.00\times1.00$ mm$^3$ \\
    NYU & Allegra 3T & 3D MPRAGE & 7$^{\circ}$ & 3.25 ms & 2530 ms & 1100 ms & None & $1.30\times1.00\times1.30$ mm$^3$ \\
    OHSU & Trio 3T & 3D MPRAGE & 10$^{\circ}$ & 3.58 ms & 2300 ms & 900 ms & None & $1.00\times1.00\times1.10$ mm$^3$ \\
    PKU 1 & Trio 3T & 3D MPRAGE & 7$^{\circ}$ & 3.39 ms & 2530 ms & 1100 ms & None & $1.30\times1.00\times1.30$ mm$^3$ \\
    PKU 2 & Trio 3T & 3D MPRAGE & 7$^{\circ}$ & 3.45 ms & 2530 ms & 1100 ms & None & $1.00\times1.00\times1.00$ mm$^3$ \\
    PKU 3 (1) & Trio 3T & 3D MPRAGE & 12$^{\circ}$ & 3.67 ms & 2000 ms & 1100 ms & None & $0.94\times0.94\times1.00$ mm$^3$ \\
    PKU 3 (2) & Trio 3T & 3D MPRAGE & 10$^{\circ}$ & 2.60 ms & 1950 ms & 900 ms & None & $1.00\times1.00\times1.30$ mm$^3$ \\
    PKU 3 (3) & Trio 3T & 3D MPRAGE  & 7$^{\circ}$ & 3.37 ms & 2530 ms & 1100 ms & None & $1.00\times1.00\times1.33$ mm$^3$ \\
    PKU 3 (4) & Trio 3T & 3D MPRAGE & 12$^{\circ}$ & 3.92 ms & 1770 ms & 1100 ms & None & $0.50\times0.50\times1.00$ mm$^3$ \\
    PKU 3 (5) & Trio 3T & 3D MPRAGE & 8$^{\circ}$ & 2.89 ms & 845 ms & 600 ms & None & $1.02\times1.02\times1.30$ mm$^3$ \\
    Pitt & Trio 3T & 3D MPRAGE & 8$^{\circ}$ & 3.43 ms & 2100 ms & 1050 ms & None & $1.00\times1.00\times1.00$ mm$^3$ \\
    WUSTL & Trio 3T & 3D MPRAGE & 8$^{\circ}$ & 3.08 ms & 2400 ms & 1000 ms & G $\times$2 & $1.00\times1.00\times1.00$ mm$^3$ \\
    \noalign{\vskip 1ex}  
    \hline
	\end{tabular}
\end{sidewaystable}
